Simulation of angular resolution of a new electromagnetic sampling calorimeter

### 2023-03-08 paper.tex
updates

In general, the sampling calorimeter enables cost-effective fabrication of large-scale detectors and design of flexible geometric configuration. This detector has been developed in various ways by selecting a material of absorber among many candidates such as brass~\cite{CMS:mat}, lead~\cite{CDF:mat}, tungsten~\cite{DELPHI:mat}, uranium~\cite{UGAS:mat}, and also of active medium such as scintillators~\cite{CMS:mat,CDF:mat}, gas~\cite{UGAS:mat}, solid-state detector~\cite{DELPHI:mat}, and liquid gas~\cite{LiqAR:mat}. Also, there are many topologies to pile up the materials and read out their signals, such as sandwich~\cite{KOTO:MB}, shashlik~\cite{shashlik:con}, spaghetti~\cite{KLOE:con}, and accordian~\cite{LiqAR:mat}. Sampling calorimeters have been used to measure the direction of arriving photons in a space-based detector to identify astrophysical sources for gamma-ray astrophysics. The PAMELA experiment employs a strip multilayer electromagnetic calorimeter made of 22 ($x$, $y$) silicon strip detectors interleaved with 22 tungsten plates, which identifies the particle types, measures the deposited energy, and helps to determine the direction of the particles~\cite{PAMELA}. The Fermi Large Area Telescope (LAT) has a tracker consisting of 18 $x$-$y$ pairs of silicon strip detectors with 16 interleaved tungsten foils. Tracking with the silicon detectors on the electron-positron pairs produced at the foils provides an angular resolution of 0.6 degrees for 1-GeV photons and better than 0.16 degrees for photons with energies higher than 10~GeV~\cite{FERMI:LAT}. The LAT has two additional components: a plastic scintillator for the charged-particle detection and a homogeneous CsI crystals for the energy measurement. This triple detector configuration is also used in different missions. An energy spectrum measurement of cosmic-ray electrons (CALET)~\cite{CALET} utilizes scintillating fibers and PWO crystals instead of the silicon strips and the CsI crystals, respectively. A dark matter particle explorer mission (DMPE)~\cite{DMPE} selects BGO crystals for the energy measurement. Theses detectors provide the angular resolution of 1-2 degrees for 1-GeV photons.

@@

The KOTO experiment at J-PARC~\cite{KOTOexp} takes another approach to identify $K_{L} \rightarrow \pi^{0}\nu\bar{\nu}$ decay with the help of a well-collimated and small-size neutral beam, constraining the $\pi^{0}$ decays along the beam axis. With this assumption, the decay vertex is calculated using the positions and energies of the two photons measured in the calorimeter. The transverse momentum of $\pi^{0}$ is calculated from the vertex. The two variables, the decay vertex and the transverse momentum are used to define a signal region of the decay. The KOTO experiment started data taking in 2013 and has achieved an experimental sensitivity of 7.2~$\times$~10$^{-10}$~\cite{KOTOexp}. The data taking will continue for a few more years, and the sensitivity is expected to reach $\mathcal{O}$(10$^{-11}$), close to the detection limit for a branching fraction near 3~$\times$~10$^{-11}$

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.58\textwidth]{Fig5_width_including7.5_12.5mm.pdf}
\caption{ Angular resolutions are deduced in terms of the strip width for 1-GeV photons at $\theta=$~10$^{\circ}$ }
\label{fig:angle_reco_width}
\end{figure}
\subsubsection{Strip width}
To inspect the dependency of the angular resolution on the strip width, we deduce the angular resolution by varying the strip width from 5 to 35mm, as shown in Figure~\ref{fig:angle_reco_width}, where each point is obtained through the fit procedure. The narrower the strip width, the larger the feature dataset size. Consequently, the angle-reconstruction performance is better with 10-mm-wide strips than with 5-mm-wide strips. However, the wider strips hardly provide detailed information on the EM shower. The 15-mm-wide strips yield an angular resolution of 1.24(1)~$\pm$~0.01$^{\circ}$ for photons impacting at the center of the strips and are selected for further study considering the actual construction of a prototype detector, although 10-mm-wide strips provide the best resolution.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.6\textwidth]{Fig8_nsample_400k_500k.pdf}
\caption{ The angular resolution for 1-GeV photons at $\theta=$~10$^{\circ}$ as a function of the number of training samples with different strip widths obtained using the first 32 layers or the full layers. The spectra are fitted with $\sqrt{I^{2} + (D/N_{\mathrm{sample}})^{k}}$.}
\label{fig:multi-parameter}
\end{figure}

>>>>

In general, the sampling calorimeter enables cost-effective fabrication of large-scale detectors and design of flexible geometric configuration. This detector has been developed in various ways by selecting a material of absorber among many candidates such as brass~\cite{CMS:mat}, lead~\cite{CDF:mat}, tungsten~\cite{DELPHI:mat}, uranium~\cite{UGAS:mat}, and also of active medium such as scintillators~\cite{CMS:mat,CDF:mat}, gas~\cite{UGAS:mat}, solid-state detector~\cite{DELPHI:mat}, and liquid gas~\cite{LiqAR:mat}. Also, there are many topologies to pile up the materials and read out their signals, such as sandwich~\cite{KOTO:MB}, shashlik~\cite{shashlik:con}, spaghetti~\cite{KLOE:con}, and accordian~\cite{LiqAR:mat}.

Sampling calorimeters have been used to measure the direction of arriving photons in a space-based detector. The PAMELA experiment employs a strip multilayer electromagnetic calorimeter made of 22 ($x$, $y$) silicon strip detectors interleaved with 22 tungsten plates, which identifies the particle types, measures the deposited energy, and helps to determine the direction of the particles~\cite{PAMELA}. The Fermi Large Area Telescope (LAT) has a tracker consisting of 18 $x$-$y$ pairs of silicon strip detectors with 16 interleaved tungsten foils. Tracking with the silicon detectors on the electron-positron pairs produced at the foils provides an angular resolution of 0.6 degrees for 1-GeV photons and better than 0.16 degrees for photons with energies higher than 10~GeV~\cite{FERMI:LAT}. The LAT has two additional components: a plastic scintillator for the charged-particle detection and a homogeneous CsI crystals for the energy measurement. This triple detector configuration is also used in different missions. An energy spectrum measurement of cosmic-ray electrons (CALET)~\cite{CALET} utilizes scintillating fibers and PWO crystals instead of the silicon strips and the CsI crystals, respectively. A dark matter particle explorer mission (DMPE)~\cite{DMPE} selects BGO crystals for the energy measurement. Theses detectors provide the angular resolution of 1-2 degrees for 1-GeV photons.

@@

The KOTO experiment at J-PARC~\cite{KOTOexp} takes another approach to identify $K_{L} \rightarrow \pi^{0}\nu\bar{\nu}$ decay with the help of a well-collimated and small-size neutral beam, constraining the $\pi^{0}$ decays along the beam axis. With this assumption, the decay vertex is calculated using the positions and energies of the two photons measured in the calorimeter. The transverse momentum of $\pi^{0}$ is calculated from the vertex. The two variables, the decay vertex and the transverse momentum are used to define a signal region of the decay. The KOTO experiment started data taking in 2013 and has achieved an experimental sensitivity of 7.2~$\times$~10$^{-10}$~\cite{KOTOexp}. The data taking will continue for a few more years, and the sensitivity is expected to reach $\mathcal{O}$(10$^{-11}$), close to the detection limit for a branching fraction near 3~$\times$~10$^{-11}$.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.58\textwidth]{Fig5_width_including7.5_12.5mm.pdf}
\caption{ Angular resolutions as a function of the strip width for 1-GeV photons at $\theta=$~10$^{\circ}$ }
\label{fig:angle_reco_width}
\end{figure}
\subsubsection{Strip width}
To inspect the dependency of the angular resolution on the strip width, we deduce the angular resolution by varying the strip width from 5 to 35mm, as shown in Figure~\ref{fig:angle_reco_width}, where each point is obtained through the fit procedure. The narrower the strip width, the larger the feature dataset size. Consequently, the angle-reconstruction performance is better with 10-mm-wide strips than with 5-mm-wide strips. However, the wider strips hardly provide detailed information on the EM shower. The 15-mm-wide strips yield an angular resolution of 1.24(1)~$\pm$~0.01$^{\circ}$ for photons impacting at the center of the strips and are selected for further study considering the actual construction of a prototype detector, although 10-mm-wide strips provide the best resolution.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.6\textwidth]{Fig8_nsample_400k_500k.pdf}
\caption{ The angular resolution for 1-GeV photons at $\theta=$~10$^{\circ}$ as a function of the number of training samples with different strip widths obtained using the first 32 layers or the full layers. The spectra are fitted with $\sqrt{I^{2} + (D/N_{\mathrm{sample}})^{k}}$.The vertical dashed-dotted line indicates the 100k events.}
\label{fig:multi-parameter}
\end{figure}
